# Data Operation

## Basic Data Type

There is 4 basic datatype in R which is:

1.  Integer
2.  Numeric or Float
3.  Character or String
4.  Logical or Boolean

**Integer** is data that has form of whole number like -10, 10, and 0. **Numeric** or float is numerical data that hasi decimal like -1.5, 2,6 and etc. **Character** or string is data type that store text like apple, fruit, etc. it is created by using quotation mark "" like this "apple". **Logical** or Boolean data type that only has two possible value which is TRUE and FALSE. Here is an example for creating each data type:

```{r}
# <- operator is used to assign value to variable
# Note: R is case sensitive so be careful when naming variable and assigning them value

# Integer
a <- 1L
## L is for whole number, we could write othe like 3L, 4L and so on

# Numeric
b <- 2.5 

# Character
c <- "stat" 

# Logical
d <- TRUE 
## Other Boolean Value is False, we could also shorthen them using T and F for True and false

# Note: Use class() to Check data type 


```

The code above means we create variable and assign them specific value using \<- operator. Here is the result of class for each variable we created

```{r}
#| echo: false
cat(paste0("Data Type of a: ", class(a), "\n",
           "Data Type of b: ", class(b), "\n",
           "Data Type of c: ", class(c), "\n",
           "Data Type of d: ", class(d), "\n"))

```

We could see for each variable has the correct data type that we has assign.

## Basic Operation

Basic Operation which is

| **Operation**  | **Symbol** | **Example** |
|:--------------:|:----------:|:-----------:|
|    Addition    |     \+     |     x+y     |
|  Subtraction   |     \-     |     x-y     |
| Multiplication |     \*     |    x\*y     |
|    Division    |     /      |     x/y     |
|    Exponent    |     \^     |    x\^y     |
|    Modulus     |     %%     |    x%%y     |

Here is an example for each basic operation

```{r}
# Define variable
a <- 10
b <- 3
c <- 2
# Addition 10+3 = 13
a+b
# Subtraction 10-3 = 7
a-b
# Multiplication 10*3 = 30
a*b
# Division 10/2 = 5
a/c
# Exponent 10^2 = 100
a^c
# Modulus 10%%2 = 0
a %% c
```

```{r}

```

## Vector

Vector is a collection of item that has the same data type or we could say it as one dimensional array. Vector in r can be created using c, here is an example
```{r}
# Creating vector
x <- c(1,2,3,4,5)
x
```
we could see the output is

## Array

```{r}
# Create an array
arr <- array(1:12, dim = c(3, 4))
# Print the array
arr
# Access an element in the array
arr[2, 3]

```


## Matrix

```{r}
# Create a matrix
mat <- matrix(1:9, nrow = 3, ncol = 3)
# Print the matrix
mat
# Access an element in the matrix
mat[2, 3]

```


## Dataframe

```{r}
# Create a dataframe
df <- data.frame(
  Name = c("Alice", "Bob", "Charlie"),
  Age = c(25, 30, 35),
  Score = c(85, 90, 95)
)
# Print the dataframe
df
# Access a column in the dataframe
df$Name
# Access a specific element
df[2, "Score"]

```


## List

```{r}
# Create a list
lst <- list(
  Name = "Alice",
  Age = 25,
  Scores = c(85, 90, 95)
)
# Print the list
lst
# Access an element in the list
lst$Scores


```
